---
title: "Downloading the occurrence data from eurOBIS"
author: "Jo-Hannes Nowé"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading the required packages
```{r packages}
library(eurobis)
library(maptools)
library(dplyr)
library(sf)
library(maps)
library(ggplot2)
library(lubridate)
library(CoordinateCleaner)
library(arulesViz)
library(tidyverse)
library(worrms)
#library("devtools")
#devtools::install_github("vlizBE/imis")
require(imis)
```


# Defining the study area
The proposed study area concerns OSPAR regions II, III 
More information about these regions can be found on: https://www.ospar.org/convention/the-north-east-atlantic.
There are observations in OSPAR regions I and IV, however when working at a monthly
resolution, the spatial extent of the points is rather limited and these two OSPAR regions
lack points in multiple year_months. 

```{r shapefiles}
#Download and unzip the OSPAR shapefiles
#Found on https://odims.ospar.org/en/submissions/ospar_regions_2017_01_002/
url <- "https://odims.ospar.org/public/submissions/ospar_regions/regions/2017-01/002/ospar_regions_2017_01_002-gis.zip"
download.file(url,paste0(spatdir,"/OSPAR_REGIONS.zip"),mode="wb")
unzip(zipfile=paste0(spatdir,"/OSPAR_REGIONS.zip"),exdir=spatdir)

#Visualize the different regions
#Need to load arulesviz package, otherwise error
OSPAR<- st_read(paste0(spatdir,"/ospar_regions_2017_01_002.shp"))
ggplot(OSPAR)+
  geom_sf()
```


As said before we only retain region II and III in order to not interpolate
our predictions to the other areas based on the few occurrences. 


```{r OSPAR}
#Only keeping region II and III from the shapefile.
OSPAR <- OSPAR[OSPAR$Region=="II"|OSPAR$Region=='III',]
#Only keeping the region and geometry info
OSPAR<- OSPAR %>% dplyr::select(Region)
#Not necessary anymore
OSPAR <- st_make_valid(OSPAR)

#Bring together the different OSPAR regions into one area
study_area <- st_union(OSPAR)
plot(study_area)
```

# Download Occurrence data from eurOBIS

```{r download-eurobis}
#Does not work on this version, but on the R3.6.3 version
#scientificname <- "Phocoena phocoena"
#aphiaid <- eurobis_name2id(scientificname)

#In this version (R4.2.2) need to directly specify the aphiaid
aphiaid <- 137117

#Does not work when knitting, but works when running seperately
mydata.eurobis<- eurobis_occurrences_full(aphiaid=aphiaid)
#save(mydata.eurobis, file=paste0("data/raw_data/mydata.eurobis.RData"))
#load(paste0(occDir,"data/raw_data/mydata.eurobis.RData"))

ggplot(study_area)+
  geom_sf()+
  geom_point(data=mydata.eurobis,x=mydata.eurobis$decimallongitude,y=mydata.eurobis$decimallatitude)
```
Before filtering the columns of interest and doing some feature engineering we create the
metadatalist, giving information about the different datasets used for the presence data.

```{r metadata-function}
# function to read dataset characteristics

fdr2<-function(dasid){
  datasetrecords <- datasets(dasid)
  dascitations <- getdascitations(datasetrecords)
  if(nrow(dascitations)==0)dascitations<-tibble(dasid=as.character(dasid),title="",citation="")
  if(nrow(dascitations)==1) if(is.na(dascitations$citation)) dascitations$citation<-""
  daskeywords <- getdaskeywords(datasetrecords)
  if(nrow(daskeywords)==0)daskeywords<-tibble(dasid=as.character(dasid),title="",keyword="")
  if(nrow(daskeywords)==1) if(is.na(daskeywords$keyword))daskeywords$keyword<-""
  dascontacts <- getdascontacts(datasetrecords)
  if(nrow(dascontacts)==0)dascontacts<-tibble(dasid=as.character(dasid),title="",contact="")
  if(nrow(dascontacts)==1) if(is.na(dascontacts$contact))dascontacts$contact<-""
  dastheme <- getdasthemes(datasetrecords)
  if(nrow(dastheme)==0)dastheme<-tibble(dasid=as.character(dasid),title="",theme="")
  if(nrow(dastheme)==1) if(is.na(dastheme$theme))dastheme$theme<-""
  dastheme2 <- aggregate(theme ~ dasid, data = dastheme, paste, 
                         collapse = " , ")
  daskeywords2 <- aggregate(keyword ~ dasid, data = daskeywords, 
                            paste, collapse = " , ")
  dascontacts2 <- aggregate(contact ~ dasid, data = dascontacts, 
                            paste, collapse = " , ")
  output <- dascitations %>% left_join(dascontacts2, by = "dasid") %>% 
    left_join(dastheme2, by = "dasid") %>% left_join(daskeywords2, 
                                                     by = "dasid")
  return(output)
}

```

```{r create-metadata}
datasetidsoi <- mydata.eurobis %>% distinct(datasetid) %>% 
  mutate(datasetid = as.numeric(str_extract(datasetid, "\\d+")))
#==== retrieve data by dataset ==============
all_info <- data.frame()
for (i in datasetidsoi$datasetid){
  dataset_info <- fdr2(i)
  all_info <- rbind(all_info, dataset_info)
}
names(all_info)[1]<-"datasetid"
write.csv(all_info,file=file.path(datadir,"allDatasets.csv"),row.names = F)
alldataset <- read.csv(file.path(datadir,"allDatasets.csv"))
```


```{r pre-processing}
# Select columns of interest
mydata.eurobis <- mydata.eurobis %>%
  dplyr::select(scientificnameaccepted,decimallongitude,decimallatitude,datecollected)%>%
  filter(!is.na(datecollected))%>%
  # Give date format to eventDate and fill out month and year columns and assign 1 to occurrenceStatus
  dplyr::mutate(occurrenceStatus = 1,day = day(datecollected),month = month(datecollected),year=year(datecollected))

#check which occurrence points fall within the study area
within_area <- st_contains(study_area,mydata.eurobis)[[1]]

#Only retain points inside the study area
mydata.eurobis <- mydata.eurobis[within_area,]

#Check if it worked as intended
ggplot(data=mydata.eurobis,x=decimallongitude,y=decimallatitude)+
  geom_sf(data=study_area)+
  geom_sf()

mydata.eurobis <- mydata.eurobis%>%
  dplyr::filter(year>=1999 & year<=2022)%>%
  arrange(datecollected)%>%
  mutate(year_month=paste(year,month,sep='-'))%>%
  mutate(year_month=factor(year_month,levels=unique(year_month),ordered=TRUE))
```


Observations already go through a couple of data quality controls before entering
the (eur)OBIS datasets. The biggest remaining issue are duplicates. These can be 
removed by using the **Coordinatecleaner** package. We consider distinct observations
with the same longitude, latitude, species and date as duplicates. Only the first
observation of each duplicate is kept in the dataset.


```{r remove-duplicates}
#Remove duplicates
mydata.eurobis <- cc_dupl(mydata.eurobis, lon = "decimallongitude", lat = "decimallatitude",
                        value = "clean",species="scientificnameaccepted", additions=
                 "datecollected")
```
# First exploratory analysis of the observations.
## Checking the temporal coverage.

# Checking spatial coverage and bias


```{r}
mydata.eurobis
save(mydata.eurobis, file = file.path(datadir,"presence.RData"))
save(study_area, file = file.path(datadir,"study_area.RData"))
```

