---
title: "Extract the environmental data associated to the species distribution data"
author: "Jo-Hannes Nowé"
output: 
html_document:
  toc=TRUE
---

# Loading the required packages
```{r}
library(terra)
```
```{r}
#Read in the PA data
load(file.path(datadir,'PA.RData'))

#Load in environmental layers
#To get it into a SpatRaster object
tempsal <- terra::rast('data/raw_data/cmems_mod_glo_phy_my_0.083deg_P1M-m_so-thetao_12.00W-13.58E_47.25N-61.75N_0.49m_1999-01-01-2021-06-01.nc')

#if the object is a spatraster you need to give a list of the length of the layers describing where you want each layer to go
splitlist <- c(rep(1,270),rep(2,270))
tempsal_dataset <- split(tempsal,splitlist)
sal <- tempsal_dataset[[1]]
sal
temp <- tempsal_dataset[[2]]
temp
```


```{r}
#This code is written to give out year_month combinations for the entirety of the environmental layers as we want to work on monthly resolution
# Generate year and month sequences
years <- 1999:2022
months <- 1:12

# Create empty vector to store year_month values
year_month_factor <- character(length(years) * length(months))

# Populate year_month vector
index <- 1
for (year in years) {
  for (month in months) {
    year_month_factor[index] <- paste0(year, "_", sprintf("%02d", month))
    index <- index + 1
  }
}

# Convert year_month to ordered factor
year_month_factor <- factor(year_month_factor, levels = unique(year_month_factor), ordered = TRUE)
PA <- PA%>%
  mutate(year_month=paste(year, sprintf("%02d", month), sep = "_"))%>%
  mutate(year_month=factor(year_month, levels = levels(year_month_factor)))
```




```{r}
#temp has 270 layers= 22*12months + 6 months (1999-01-01 till 2021-06-01)
PA <- PA%>%filter(datecollected<"2021-07-01")
```

```{r}
#NPP
npp <- terra::rast("data/raw_data/cmems_mod_glo_bgc_my_0.083deg-lmtl_PT1D-i_npp_12.00W-13.58E_47.25N-61.75N_1999-01-01-2022-12-31.nc")
npp
tijd <- time(npp)
ym_tijd <- format(tijd, "%Y_%m")
ym_tijd <- factor(ym_tijd, levels = unique(ym_tijd), ordered = TRUE)
splitlijst <- as.numeric(ym_tijd)
npp_dataset <- split(npp,splitlijst)
npp_dataset
mean_npp <- list()
for(i in 1:length(npp_dataset)){
  mean_npp[[i]] <- app(npp_dataset[[i]],fun=mean)
}
time(mean_npp)
mean_npp <-rast(mean_npp)
```


#Now for the bathymetry layer

Manually create the folder "bathy_sliced" first, tried to include it in the code but for some reason couldn't create it.

```{r}
## general data handling
library(XML)
library(dplyr)
library(tidyr)
library(reshape2)
library(downloader)

library(raster)
library(sp)
library(mapdata)
library(maptools)
library(ncdf4)
library(tiff)

ifelse(!dir.exists(file.path(envdir, "bathy_sliced")), dir.create(file.path(envdir, "bathy_sliced")), FALSE)
xmin<- -12.0
xmax <- 13.6
ymin <- 47.2
ymax <- 61.8
stepsize <- 0.8
number_of_slices <- (xmax-xmin)/stepsize
for(i in 1:number_of_slices){
beginslice <- xmin + (i-1)*stepsize
endslice <- xmin + i*stepsize
#We get an error because the file is too big, will split up in horizontal slices, and then merge together later 
con <-paste0("https://ows.emodnet-bathymetry.eu/wcs?service=wcs&version=1.0.0&request=getcoverage&coverage=emodnet:mean&crs=EPSG:4326&BBOX=",beginslice,",47.2,",endslice,",61.8&format=image/tiff&interpolation=nearest&resx=0.08333333&resy=0.08333333")
nomfich <- file.path(envdir,paste0("bathy_sliced/","slice_",i, "_img.tiff"))
download(con, nomfich, quiet = TRUE, mode = "wb")

}


#merge them together
bathyrasters <- list()
for(file in list.files(file.path(envdir,"bathy_sliced"))){
 bathyrasters[[file]] <- rast(file.path(envdir,paste0("bathy_sliced/",file))) 
}
#Put the different spatrasters together in a spatrastercollection
bathy_coll <- sprc(bathyrasters)

#Merge the spatrastercollection
bathy <- merge(bathy_coll)
bathy[bathy>0] <- NA
```

# Extraction of values related to the occurrence points

Now that we have the environmental layers loaded as spatial rasters, we need to extract the right values.
It is easiest to download them first for all the months and then just select the right monthly value for each point.

```{r}
#For temperature
rastertemp <- terra::extract(x=temp, y=PA[,c("decimallongitude","decimallatitude")], method="bilinear", na.rm=TRUE,df=T,ID=FALSE)
#takes 20seconds
resulttemp <- data.frame(temp = rastertemp[cbind(1:nrow(PA),as.numeric(PA$year_month))])

#For NPP
rasternpp <- terra::extract(x=mean_npp, y=PA[,c("decimallongitude","decimallatitude")], method="bilinear", na.rm=TRUE,df=T,ID=FALSE)
#takes 20seconds
resultnpp <- data.frame(npp = rasternpp[cbind(1:nrow(PA),as.numeric(PA$year_month))])


#For salinity
rastersal <- terra::extract(x=sal, y=PA[,c("decimallongitude","decimallatitude")], method="bilinear", na.rm=TRUE,df=T,ID=FALSE)
#takes 20seconds

resultsal <- data.frame(sal = rastersal[cbind(1:nrow(PA),as.numeric(PA$year_month))])

#For bathymetry
resultbathy <- terra::extract(x=bathy, y=PA[,c("decimallongitude","decimallatitude")], method="bilinear", na.rm=TRUE,df=T,ID=FALSE)
#don't need to choose the right month, because bathy info is the same over the different months, only have one value

```

```{r}
ggplot()+
  geom_spatraster(data=bathy)

mean_npp_test <- mean_npp[[1]]

ggplot()+
  geom_spatraster(data=mean_npp[[1]])

ggplot()+
  geom_spatraster(data=temp[[1]])

ggplot()+
  geom_spatraster(data=sal[[1]])
```

```{r}
PA_env <- cbind(PA,resulttemp,resultsal,resultnpp,resultbathy)
save(PA_env,file=file.path("data","PA_env.RData"))
```


