---
title: "Tidymodels ensemble approach"
author: "Jo-Hannes Nowé"
output: 
html_document:
  toc=TRUE
---

# Loading the required packages
```{r}
library(tidymodels)
#devtools::install_github("jiho/autoplot")
library(autoplot)
library(stacks) #for ensembling the individual models together
library(workflows)
library(ranger) #for the randomforest model
library(mgcv) #for the GAM model
library(xgboost) #for the xgboost model
library(earth) #for the MARS model
```

# Load data

We start with presence-(pseudo)absence data that is coupled to certain environmental layers.
This is the necessary starting point of this script, from here on we then build the 
different individual models and combine them in an ensemble. (Possible to run the different
models in parallel?).

```{r}
# Load presence-absence related to environmental data
PAdata <- file.path(datadir,"PA_env.RData")

load(PAdata)
#Turn our occurrence into a factor so that there are two known levels, either
#present or absent.
PA_env$occurrenceStatus<-factor(PA_env$occurrenceStatus)
#Remove all these columns because the data we want to predict on also only has the environmental values,
#model will only accept input to predict on with the same columns as the data the model was trained on.
PA_env <- PA_env%>%select(-c(scientificnameaccepted,decimallongitude,decimallatitude,datecollected,day,month,year,year_month,geometry))%>%
  drop_na()%>%
  mutate(bathy=slice_1_img)%>%
  select(-slice_1_img)

#Why make this into a tibble?
data <- as_tibble(PA_env)%>%
  dplyr::select(c(-ID,-scientificname))
data
```

# Data splitting
Need to split into a test and training set. Training set is used for training and validation.
The test set is kept separate and only used for performance evaluation.
We use a 80/20 distribution of train-test.
Out of this 80% training set, we use 10% each fold for validation.

Still need to implement cross-validation

```{r}
set.seed(222)
#Put 4/5 in the training set
data_split<- initial_split(data, prop=4/5)

#Create data frames for the two sets:
train_data <- training(data_split)
test_data <- testing(data_split)

#Create folds in the training data so we can do our validation.
#Out of the 80% reserved for training data, use 10% each time for validation
#This is why we choose the number of folds v=8
folds <- vfold_cv(train_data, v=8)
folds
```


# Pre-process your data with recipes
Recipes can be used to define the formula, the dependent variable, predictors, others.
The roles of the recipe can also be updated when necessary. This recipe is then 
provided to the workflow, giving information about the necessary relations to calculate.

Variables are the original columns in the data. These can then be used to define a
formula. Roles define how variables will be used in the model. Examples are predictor,
response, and case weight. Terms are columns in a design matrix, synonymous with
features in machine learning. Variables that have predictor roles would automatically be
main effect terms. 

```{r}
#The general form is: recipe(formula, data)

#The formula is only used to declare the variables, their roles and nothing else.
#All the rest can be added later.

#Doesn't matter the data here is train_data. Just used to catalog the names of the 
#variables and their types. 
#The formula here states that occurrenceStatus is modelled in relation to all the other columns
Occurrence_rec <- 
  recipe(occurrenceStatus ~., data=train_data)

#Can also do some pre-processing of the variables with this recipe, but which should you do?
# step_* functions
#Can also provide some checks here. These check_* functions conduct some sort of 
#data validation, if no issue is found, they return the data as it is, otherwise they show an error.
```


The dot indicates to use all other variables as predictors
A recipe is associated with the data set used to create the model. We can add roles to the recipe, using
update_role() lets the recipe know that two of the variables have the role "Location"
so it tells the recipe to keep these two variables but not use them as either outcomes
or predictors. Keeping some variables without using them in the model can be useful
for determining if certain values were wrongly predicted. 

# Fit a model with a recipe, using a model workflow
# Making an ensemble model with stacks

In stacks, you need to save the assessment of predictions and workflow in your
tune_grid(), tune_bayes() or fit_resamples() objects by setting control arguments
save_pred=TRUE and save_workflow = TRUE. 

Each model definition must share the same rsample rset object. 

```{r}
#We already generated our our rsample rset objects
data_split
train_data
test_data

#We also already made some folds
folds

#And a recipe
Occurrence_rec

#Define a certain metric?
#metric <- metric_set(rmse)
```
Above we talked about saving the information in the resulting objects. This can also
be done by using control_stack_*() functions.

```{r}
#Because we use tune_grid()
ctrl_grid <- control_stack_grid()
ctrl_res <- control_stack_resamples()
```

The random forest model is a model that does not require a lot of tuning, which is 
why we just give fixed values.
```{r}

#set the number of trees to 500 (also the default), use it for classification into
#presence-absence, and the algorithm to ranger (also the default)
rf_mod <- rand_forest(trees=500,mode="classification",engine="ranger")

#After your model is defined, create a worfklow object and add your model and recipe
rf_wf <-
  workflow() %>%
  add_model(rf_mod) %>%
  add_recipe(Occurrence_rec)

#The workflow can then be fit or tuned, in this case it's fit but with resamples
#so we have to provide the folds
#Runs +- 2min
rf_fit <-
    fit_resamples(
    rf_wf,
    resamples = folds,
    control = ctrl_res
  )
#With resamples as we need to provide this type to the stack command later
rf_fit
```

```{r}
rf_fit
#Predict based on the test set, as noted need to make folds of this
#predict and ppass the type parameter to get out a probability.
rf_pred <- predict(rf_fit,test_data)%>%
  bind_cols(predict(rf_fit,test_data, type="prob")) %>%
  #To be able to compare the real value to the predicted one
  bind_cols(test_data %>%
              dplyr::select(occurrenceStatus)) #need to specify due to conflict by loading tidymodels
rf_pred
#Just one of the possible evaluation metrics, need to decide which ones to use
rf_pred %>%
  roc_auc(truth = occurrenceStatus, .pred_0)
  
rf_pred %>%
  accuracy(truth= occurrenceStatus, .pred_class)

```

# Tuning model parameters
```{r}
#More information found on https://parsnip.tidymodels.org/reference/details_gen_additive_mod_mgcv.html
gam_mod <- gen_additive_mod(adjust_deg_free = tune()) %>% 
  set_engine("mgcv") %>% 
  set_mode("classification")
gam_wf <-
  workflow()%>%
  add_recipe(Occurrence_rec)%>%
  add_model(gam_mod, formula = occurrenceStatus ~ s(bathy)+ s(temp)+ s(sal) + s(npp))
gam_grid <- grid_regular(adjust_deg_free(),
                         levels=4)
gam_fit <-
  gam_wf%>%
  tune_grid(
    resamples = folds,
    grid=gam_grid,
    control=ctrl_grid
  )
#started at 16:30 - 16:43
# gam_res <-gam_fit%>%
#   collect_metrics()
# gam_best <- gam_fit%>%
#   select_best("accuracy")
# gam_final_wf <- gam_wf%>%
#   finalize_workflow(gam_best)
# #Need to also give in the control here, otherwise doesn't work
# gam_fit <- gam_final_wf%>%
#   last_fit(data_split,
#            control=ctrl_grid)
```

```{r}
#Tune the xgboost model, see: https://parsnip.tidymodels.org/reference/details_boost_tree_xgboost.html
xgb_mod <- boost_tree(
  trees = tune(),
  tree_depth = tune(),
  learn_rate = tune()
) %>%
  set_engine("xgboost") %>%
  set_mode("classification")
xgb_wf <- workflow() %>%
  add_model(xgb_mod) %>%
  add_recipe(Occurrence_rec)
#grid_regular chooses sensible hyperparameter values to try, but want to choose ourself
#Can give th elower and upper bound limit we want to use
xgb_grid <- parameters(trees(),
                          tree_depth(c(3,20)),
                          learn_rate(c(-4,0)))
xgb_grid <- grid_regular(xgb_grid,
                           levels=c(trees=5,tree_depth=3,learn_rate=3))
```

```{r}
xgb_fit <- 
  xgb_wf %>% 
  tune_grid(
    resamples = folds,
    grid = xgb_grid,
    control=ctrl_grid)
#Runs for +- 40min
xgb_fit
```

```{r}
#devtools::install_github("tidymodels/tune")
#Need to update to this version of the tune package, otherwise it fails
mars_mod <- 
    mars(prod_degree = tune(), prune_method = tune()) %>% 
    # This model can be used for classification or regression, so set mode
    set_mode("classification") %>% 
    set_engine("earth")#,glm=list(family=binomial))
mars_mod
mars_grid <- grid_regular(prod_degree(),prune_method(),
                          levels=c(prod_degree=2,prune_method=2))
mars_grid
mars_wf <-
  workflow()%>%
  add_model(mars_mod)%>%
  add_recipe(Occurrence_rec)
mars_fit <-
  mars_wf%>%
  tune_grid(
    resamples = folds,
    grid=mars_grid,
    control=ctrl_grid
  )

mars_fit
```

Now specify a random forest model that we are not optimizing.


Now we have XX candidate models (1 RF, xx GAM, xx XGB and xx MARS). 
We want to create a data stack: tibbles containing an assessment set predictions for 
each candidate ensemble member. It work a bit like the basic ggplot() constructor.
The function creates a basic structure that the object will be built on top of.

```{r}
stack_data <- 
  stacks() %>%
  add_candidates(rf_fit) %>%
  add_candidates(mars_fit)%>%
  add_candidates(gam_fit)%>%
  add_candidates(xgb_fit)
  
stack_data
```

The first column gives the first response value and the remaining columns give the
assessment set predictions for each ensemble member. In regression there is 1 column
per ensemble member. In classification settings, as many columns as levels of the outcome variable
per candidate ensemble member. 

```{r}
stack_mod <-
  stack_data %>%
  blend_predictions()
#16h28-16h29
#Problem that there are much more member models for xgboost?
#However there was only 1 random forest and seemed to be good.
stack_mod
```
Blend predictions performs regularization to figure out how we can combine the outputs
from the stack members to come up with the final solution. Regularization needed
as the ensemble members are highly correlated. Candidates with non-zero stacking 
coefficients become members. 
We can tune the penalty parameter of this blend_predictions as well, see:
https://stacks.tidymodels.org/articles/basics.html

```{r}
#Now fit the whole model on the training set
stack_fit <-
  stack_mod %>%
  fit_members()
#16h30 - 16h33
stack_fit
```


```{r}
#Saving the model
saveRDS(stack_fit,file.path(datadir,"stacked_model.rds"))

#testing if the saving went as planned
#stack_fit <- readRDS("stacked_model.rds")
```






